Utilization of Reductions and Abstraction Elimination in Typed Genetic Programming

### 2014-01-27 tex/gecco14/gecco14.tex
gecco

@@

Zdá se že použití kombinátorů má řadu výhod (vymenovat nnějaký co jmenuje briggs), 
ale konzistentní prozkoumání obecnýho řešení lambdovýho charakteru určitě stojí 
taky za zkoužku, což podporují nasledující argumenty:

@@

\Lets describe transformation $\Ae$ performing this 
process.\footnote{In our implementation we must also deal with types, because 
our individuals are annotated with types. But since this process
is fairly straightforward but cumbersome to describe, we will 
skip explaining it here. Reader interested in seeing it can see 
the source code of our implementation. }
\begin{align*}
\Ae[x]           &= x &\\[0.4em]
\Ae[\,(M\,N)\,]  &= (\Ae[M]\;\;\Ae[N]) &\\[0.4em]
\Ae[\lam{x}{x}]  &= \Icomb &\\
\Ae[\lam{x}{M}]  &= (\Kcomb~\Ae[M]) &\textbf{if } x \not\in \FV(M)\\
\Ae[\lam{x}{\lamb{y}{M}}] &= \Ae[\lam{x}{\Ae[\lam{y}{M}]}]  
&\textbf{if } x \in \FV(M)\\
\Ae[\lam{x}{(M\,N)}] &= (\Scomb~\Ae[\lam{x}{M}]~\Ae[\lam{x}{N}])  
&\textbf{if } x \in \FV(M)\\
&&\vee~ x \in \FV(N)\\
\end{align*}

@@

@@

\red{
V této sekci prodiskutujeme generování termův lnf, kteroužto můžeme chápat jako
příme zobecnění S-expressions do lambda termů. Povíme si jaký vztah takové termy mají
k beta a eta redukci, jaká je jejich výhoda a jak je lze jednoduše převést
do úspornější beta eta nf. 
}

@@

\red{řict že lnf je přirozený rozšíření s-exprešnu do lambda 
kalkulu: pokud vezmeme gama že splnuje closure podmínku, bude 
se používat jen pravidlo jedna, one can see that f stands for 
root node and (ty závorky) for direct subtrees. For context containing 
at least one higer ord function we get special kind of root node containing
lambda head with one direct subtree standing for body of 
the lambda function, v tomto tele se navíc mohou objevit proměnné 
definované v této hlavě (asi doplnit obrázkama).

@@

Since \ered only destroys and never creates \textit{function applications} (i.e. @),
the root @ of $B$ must be present in $P$.  
But since $P$ contains no \bredex, the left subterm $L$ of this root @
is not $\lambda$-abstraction.
Only possible way for $L$ to be changed by $\eArrow$ into 
a $\lambda$-abstraction is that $L$ is the reduced subterm (so that
$L$ is changed for its subterm).
But that is in contradiction with $P$ not containing any \bredex,
because it would cause $L$ be a $\lambda$-abstraction.

\end{proof}

@@

\red{Hlavní problém: proměnné - pokud se odtržený podterm obsahuje volnou proměnnou, taková proměná může ve svém novém umístění porušit well typednost vzniklého termu - ať už z důvodu toho že tam není vubec definována, nebo tam sice je, ale má jiný typ.

Metody jak zacházet s lambdama:

- použít méně obecnou reprezentaci, takovym
  příkladem je Yu, která nepovoluje vnější proměný uvnitř
  lambda abstrakcí a další restrikce. Touto cestou nechceme jít,
  myslime si že stojí za to udělat takovou reprezentaci, která nepřeskočí
  normální formu žádného well-typed termu.

- křížit tak jako by se nic nedělo a problémy řešit až když nastanou:
  tzn buď přejmenovat promenou na takovou, která v daném kontextu má správný typ nebo ji nahradit, případně ji nahradit nově vygenerovaným termem požadovaného typu. 
  Takový přístup se nam taky nelíbí, dá se očekávat že takové kolize budou častý, při nemožnosti vhodně přejmenovat proměnnou navíc do křížení přibívá
  prvek mutace. K takovému řešení se zdá dobé přistoupit až ve chvíli, kdy nezbejvá jiná čistčí možnost.

Pro vyřešení problémů s proměnými se dá použít elegantní řešení: zbavit se proměných. Toho jde docílit použitím algoritmu eliminace abstrakcí, který převede term na ekvivalentní term který neobsahuje žádné proměné a lamb abstrakce, zato obsahuje navíc kombinátory (v nejjdednoduší verzi S,K a I), které zastanou puvodní roli proměnných. We idetified two methods using this approach:

\subsection{Pokračujou zkopčenosti z théze až dokonce }

\red{todle se dá dát asi na začátek crossover operation spíš, nebo nevim}

Design goal behind our approach to crossover operation is
to try to generalize standard tree swapping crossover.
This is not because of dislike of more specialized crossovers;
it comes from systematic approach for exploring the domain of
typed functional GP --- thus I would like to start with the most common one.\\

\red{to furt platí, protože na s-výrazy se eta redukce ani abstr elim nevztahuje}

The crossover operation in standard GP is performed 
by swapping randomly selected subtrees in each parent 
S-expression.

For typed lambda terms two difficulties arise: Types and variables.\\

\red{CO S @-stromama řešit je explicitně tady, případně někde jinde? Pro generování nemá reprezentace vliv, ta má vliv až vpři křížení, kde at-tree zajišťuje víc míst. We will show how to crossover \textit{typed \lterm trees} in both \atTree and \sexprTree notation.}

As in standard GP our crossover will be performed by swapping
two subtrees. But now with constraint that both subtrees have
the same type. This is discussed in grater detail in \ref{typed-swapping}.\\

@@

% - This is only one example of a problem caused by swapping subtrees with variables,
% - Resolve problems with free variables or avoid variables completely.

%In \ref{repairing-method} is presented more detailed discussion of problems with
%variables and method which repairs such "broken" terms.

First approach tried at this problem was to repair the broken term,
but since this method is not currently implemented we will describe it
in just a few words. If the moved free variable ends in the scope 
where is defined variable with the same name and type, then it
is OK. If it is not this case, we may rename this variable to some 
other variable "visible" in its new scope if such a variable has the same 
type. If there is no such a suitable variable we may generate new value 
to replace the variable by term generating method.

As you can see, this is not much elegant solution.

But we can avoid dealing with this problem by avoiding use of variables.
This can be achieved by process called abstraction elimination.

\subsubsection{Typed subtree swapping}
\label{typed-swapping}

@@

\red{mimojiné zmínit, že v budouznu bysme chtěli porovnat kombinátory 
s tou naší nějak extenzivně (což by bylo hezký už ted ale nestíhá se to..)}

>>>>

Todele asi šoupnout do related work.

@@

Zdá se že použití kombinátorů má řadu výhod (vymenovat nějaký co jmenuje briggs), 
ale konzistentní prozkoumání obecnýho řešení lambdovýho charakteru určitě stojí 
taky za zkoužku, což podporují nasledující argumenty:

@@

\Lets describe transformation $\Ae$ performing this 
process.
%\footnote{In our implementation we must also deal with types, because our individuals are annotated with types. But since this process is fairly straightforward but cumbersome to describe, we will skip explaining it here. Reader interested in seeing it can see the source code of our implementation. }
\begin{align*}
\Ae[x]           &= x &\\[0.4em]
\Ae[\,(M\,N)\,]  &= (\Ae[M]\;\;\Ae[N]) &\\[0.4em]
\Ae[\lam{x}{x}]  &= \Icomb &\\
\Ae[\lam{x}{M}]  &= (\Kcomb~\Ae[M]) &\textbf{if } x \not\in \FV(M)\\
\Ae[\lam{x}{\lamb{y}{M}}] &= \Ae[\lam{x}{\Ae[\lam{y}{M}]}]  
&\textbf{if } x \in \FV(M)\\
\Ae[\lam{x}{(M\,N)}] &= (\Scomb~\Ae[\lam{x}{M}]~\Ae[\lam{x}{N}])  
&\textbf{if } x \in \FV(M)\\
&&\vee~ x \in \FV(N)\\
\end{align*}

@@

\red{tady je to v původnim kontextu, odebrat slova co se tam nehodí}

@@

This section is focused on the individual generating method producing
terms in their long normal form, which can be understood as a straight 
generalization of S-expressions into lambda calculus. 
We discuss the relation between terms in \lnf with beta and eta reductions,
the advantages of such representation and we also show how to easily 
transform such terms into short \benf.

@@

Long normal form is a generalization of S-expressions into lambda calculus:
Let $\Gamma$ be a context satisfying the closure requirement, then only
the first rule will be applicable. 
$(\alpha, \Gamma)$ rewrites to $(~f~( \alpha, \Gamma )~\dots~( \alpha, \Gamma )~)$.
One can see that $f$ stands for 
the root node and that each $( \rho_i , \Gamma )$ will produce a direct subtree.

\red{
For context containing at least one higher order function we get special kind of root node containing lambda head with one direct subtree standing for body of 
the lambda function, v tomto tele se navíc mohou objevit proměnné 
definované v této hlavě (asi doplnit obrázkama).

@@

Since \ered only destroys and never creates \textit{function applications} (i.e. @),
the root @ of $B$ must be present in $P$.  
But since $P$ contains no \bredex, the left subterm $L$ of this root @
is not $\lambda$-abstraction.
Only possible way for $L$ to be changed by $\eArrow$ into 
a $\lambda$-abstraction is that $L$ is the reduced subterm (so that
$L$ is changed for its subterm).
But that is in contradiction with $P$ not containing any \bredex,
because it would cause $L$ be a $\lambda$-abstraction.
\end{proof}

@@

The design goal behind our approach to the crossover operation is
to try to generalize the standard tree swapping crossover.

The crossover operation in standard GP is performed 
by swapping randomly selected subtrees in each parent 
S-expression.
For typed lambda terms two difficulties arise: Types and variables.

As in standard GP our crossover will be performed by swapping
two subtrees. But now with constraint that both subtrees have
the same type.

@@

Generally speaking, the variable problem is caused by that the local variable is not defined in the new place or the variable is defined in the new place, but has
some another type.

Let us list some possible approaches to solving this problem:

%Už přeloženo: 
%Hlavní problém: proměnné - pokud se odtržený podterm obsahuje volnou proměnnou, taková proměná může ve svém novém umístění porušit well typednost vzniklého termu - ať už z důvodu toho že tam není vubec definována, nebo tam sice je, ale má jiný typ.  
%Metody jak zacházet s lambdama:
%- použít méně obecnou reprezentaci, takovym příkladem je Yu, která nepovoluje vnější proměný uvnitř lambda abstrakcí a další restrikce. Touto cestou nechceme jít, myslime si že stojí za to udělat takovou reprezentaci, která nepřeskočí normální formu žádného well-typed termu.
%- křížit tak jako by se nic nedělo a problémy řešit až když nastanou: tzn buď přejmenovat promenou na takovou, která v daném kontextu má správný typ nebo ji nahradit, případně ji nahradit nově vygenerovaným termem požadovaného typu. Takový přístup se nam taky nelíbí, dá se očekávat že takové kolize budou častý, při nemožnosti vhodně přejmenovat proměnnou navíc do křížení přibívá prvek mutace. K takovému řešení se zdá dobé přistoupit až ve chvíli, kdy nezbejvá jiná čistčí možnost.
%Pro vyřešení problémů s proměnými se dá použít elegantní řešení: zbavit se proměných. Toho jde docílit použitím algoritmu eliminace abstrakcí, který převede term na ekvivalentní term který neobsahuje žádné proměné a lamb abstrakce, zato obsahuje navíc kombinátory (v nejjdednoduší verzi S,K a I), které zastanou puvodní roli proměnných. We idetified two methods using this approach:

One option is to choose a less general representation of lambda terms.
Such approach is successfully used in \cite{yu01} where the 
terminal set for constructing lambda abstraction subtrees 
is limited to use only constants and variables of that particular
lambda abstraction and an appropriate variable naming convention is
used in order to prevent troubles with variables.
We choose not to follow such approach because we do not want to
lose the opportunity to represent every possible well-typed 
term in its normal form.

Another option is to overlook the variable problems and to correct the
defects when they occur. Such correction can be performed by 
renaming the problematic variable to some another variable defined in that place 
or, if there is no suitable candidate, to replace its occurrence with 
a newly generated term of required type. This approach seems to us as a very
unwieldy one -- there are clearly problems for which such collisions would
occur very often and when there is no suitable rename candidate the 
generation of new term brings the unnecessary element of a mutation.
Such solution seems to be useful only as a last resort.

There is an elegant way to overcome the problem with variables by getting rid of them.
This is possible thanks to the abstraction elimination algorithm described above.
It turns a \lterm into a \lterm that contains no lambda abstractions
and no variables, instead, it contains additional function symbols standing for 
polymorphic combinators (i.e. $\Scomb$, $\Kcomb$ and $\Icomb$ 
in the simple case of the algorithm).

We identified two methods using this approach:

\red{

\subsection{Poznámky k uvodu ke křížení}
\red{
POZNAMKY:

CO S @-stromama řešit je explicitně tady, případně někde jinde? Pro generování nemá reprezentace vliv, ta má vliv až vpři křížení, kde at-tree zajišťuje víc míst. We will show how to crossover \textit{typed \lterm trees} in both \atTree and \sexprTree notation.

to o tom že chem zobecnit tu standardní furt platí, protože na s-výrazy se eta redukce ani abstr elim nevztahuje, někde to zmínit že se to na ně nevztahuje
}

\subsection{Typed subtree swapping in greater detail}
\label{typed-swapping}

@@

\red{mimojiné zmínit, že v budouznu bysme chtěli porovnat kombinátory 
s tou naší nějak extenzivně, ikdyž možná to tam vubec nepsat
(což by bylo hezký už ted ale nestíhá se to..)}
